Temporary Trade Barriers: \\ When Will They End?

### 2016-05-10 presentation.tex
strip more out of old presentation

\usepackage{cite}
\usepackage{times, verbatim,xcolor,bm}
\usepackage{amsbsy,amssymb, amsmath, amsthm}
\usepackage{booktabs}
%David miller's fonts
	\usepackage[T1]{fontenc}
	\usepackage[boldsans]{ccfonts}
	\usepackage[euler-hat-accent]{eulervm}

@@

\section{Overview}
\subsection{Preview}
\begin{frame}
\frametitle{Political Economy of Trade Agreements}
In most models, a single actor in each country makes trade policy. Modeling policy-making power more richly:
\pause
\begin{itemize}[<+->]
	\item reconciles `Protection for Sale' model to the data
  \item generates lobbying for the purpose of disrupting trade agreements  
	\item adds new role for trade agreements 
\end{itemize}
\end{frame}

\begin{frame}{Preview of Results}
\begin{itemize}[<+->]
	\item Higher trade agreement tariffs induce lobbies to decrease effort 
		\begin{itemize}
			\item Reduces probability of trade war / ratification failure
		\end{itemize}
	\item Executives negotiate tariffs to discourage lobbying
		\begin{itemize}
			\item Produces higher tariffs than executives prefer
			\item Reduces equilibrium lobbying effort
			\item Trade agreement as political commitment device
		\end{itemize}
	\item Tariff levels and optimal lobbying depend on how much political uncertainty is present
\end{itemize}
\end{frame}

\begin{frame}{Outline of Talk}
\begin{enumerate}[<+->]
	\item Background and Motivation
	\item Economic and Political Structure
	\item Results for a Special Case: Certainty
	\item Full Results: Political Uncertainty
	\item Conclusion
\end{enumerate}
\end{frame}

\begin{frame}{Standard Unitary Government Model}
Grossman $\&$ Helpman government welfare function:
				\[
				  C + aW
				\]
\pause
		\begin{itemize}[<+->]
			\item Government auctions tariffs/subsidies: lobbies pay for $a$-weighted welfare loss
			\item Estimated $a$'s far too large
				\begin{itemize}
					\item Implies that governments respond very little to lobbying effort
					\item Inconsistent with observed large tariffs relative to small lobbying effort
				\end{itemize}
		\end{itemize}
\end{frame}

\begin{frame}{Trade Agreements and Separation of Powers}
  \pause
  Take special case of `Protection for Sale' model (single lobby) and generalize government structure
  \pause
		\begin{itemize}[<+->]
			\item Executives can make trade agreement
			\item Legislatures retain final authority: can break agreement and start trade war / fail to ratify
			\item Trade war is subgame of the trade-agreement-setting game
				\begin{itemize}
  				\item Executives consider lobbying incentives and possibility of trade war when forming agreement
				\end{itemize}
		\end{itemize}
\end{frame}

\subsection{}
\begin{frame}
\frametitle{Related Literature}
\small Protection for Sale: Grossman $\&$ Helpman (1994)
\begin{itemize}
  \item \footnotesize Empirics: Goldberg $\&$ Maggi (1999), Gawande $\&$ Bandyopadhyay (2000), Mitra, Thomakos, $\&$  Ulubasoglu (2002)
  \item \footnotesize Mitra, Thomakos, $\&$  Ulubasoglu (2006), Bombardini (2008)
	\item \footnotesize Trade Wars and Trade Talks: Grossman $\&$ Helpman (1995)
\end{itemize}

@@

		%\item Autocracy vs. Democracy: Mitra, Thomakos and Ulubasoglu (2002), Aidt and Gassebner (2010)
\end{frame}

@@

\vskip.2in
\begin{itemize}
	\item $e_b$: Lobbying effort for break decision
	\item $e_{tw}$: Lobbying effort for trade-war tariffs
	\item $\bm{\tau^a} = (\tau^a,\tau^{*a})$: tariff under an agreement
	\item $\bm{\tau^{\mathit{tw}}} = (\tau^{\mathit{tw}},\tau^{*\mathit{tw}})$: tariff when agreement is broken
\end{itemize}

@@

\section{Certainty}
\subsection{Results: The Certain Case}
\begin{frame}{Timeline}
\begin{enumerate}
	\item \textbf{Formation}
		\begin{enumerate}[i.]
			\item Executives set trade policy in international agreement
		\end{enumerate}
	\item \textbf{Ratification / Maintenance}
		\begin{enumerate}[i.]
			\item Firms lobby legislatures to break agreement
			\item Legislatures decide to break or abide by agreement
		\end{enumerate}
	\item {\color{gray} \textbf{Trade War} (if agreement is broken)}
		\begin{enumerate}[i.]
			\item {\color{gray} Firms lobby legislatures to set high trade-war tariff}
			\item {\color{gray} Legislatures decide trade-war tariff}
		\end{enumerate}
	\item {\color{gray} Private actors make production, consumption decisions}
\end{enumerate}
\end{frame}

@@

\section{Full Results}
\begin{frame}{Political Uncertainty Illustration}
\pause
An Example (Bagwell $\&$ Staiger 2005)
\pause
\begin{itemize}[<+->]
	\item $D(P_i) = 1 - P_i$
	\item $Q_X(P_X) = \frac{P_X}{2}$, $Q_Y(P_Y) = P_Y$
	\item $P_X^W = \frac{4-3\tau}{7}$, $P_X = \frac{4+4\tau}{7}$
	\item $\ga(e,\ta) = 1.25 + e^{0.2} + \ta$ 
		\begin{itemize}
			\item $\ta \sim U[-0.25,0.25]$
		\end{itemize}
	\item $\ga_E = 1$
\end{itemize}

@@

\subsection{Ratification / Maintenance Stage}
\begin{frame}{Timeline}
\begin{enumerate}
	\item {\color{gray} \textbf{Formation}}
		\begin{enumerate}[i.]
			\item {\color{gray} Executives set trade policy in international agreement}
		\end{enumerate}
	\item \textbf{Ratification / Maintenance}
		\begin{enumerate}[i.]
			\item {\color{gray} Firms lobby legislatures to break agreement}
			\item Legislatures decide to break or abide by agreement
		\end{enumerate}
	\item {\color{gray} \textbf{Trade War} (if agreement is broken)}
		\begin{enumerate}[i.]
			\item {\color{gray} Firms lobby legislatures to set high trade-war tariff}
			\item {\color{gray} Legislatures decide trade-war tariff}
		\end{enumerate}
	\item {\color{gray} Private actors make production, consumption decisions}
\end{enumerate}
\end{frame}

@@

\subsection{Formation Stage}
\begin{frame}{Timeline}
\begin{enumerate}
	\item \textbf{Formation}
		\begin{enumerate}[i.]
			\item Executives set trade policy in international agreement
		\end{enumerate}
	\item {\color{gray} \textbf{Ratification / Maintenance}}
		\begin{enumerate}[i.]
			\item {\color{gray} Firms lobby legislatures to break agreement}
			\item {\color{gray} Legislatures decide to break or abide by agreement}
		\end{enumerate}
	\item {\color{gray} \textbf{Trade War} (if agreement is broken)}
		\begin{enumerate}[i.]
			\item {\color{gray} Firms lobby legislatures to set high trade-war tariff}
			\item {\color{gray} Legislatures decide trade-war tariff}
		\end{enumerate}
	\item {\color{gray} Private actors make production, consumption decisions}
\end{enumerate}
\end{frame}

\begin{frame}{Total Break Probability}
$B(\bta) = b(e_b(\bta),\bta)$
\pause
\begin{itemize}[<+->]
	\item Both direct and indirect effects of raising $\bta$ are negative
		\begin{itemize}
			\item $B(\bta)$ decreasing in $\bta$
		\end{itemize}
\end{itemize}
\pause
\vskip.2in
\begin{beamerboxesrounded}[upper=palette tertiary, shadow=true]{Result 3}
   The total probability that the trade agreement will be broken is decreasing in the trade agreement tariffs.
\end{beamerboxesrounded}

\begin{frame}{Executives' Tradeoff}
\[
  \Pr\left[ \text{TradeWar}(\bta) \right] \bm{W_E}(\btn) + \left\{1- \Pr\left[ \text{TradeWar}(\bta) \right]\right\} \bm{W_E}(\bta)
\]
\pause
\vskip.2in
\begin{beamerboxesrounded}[upper=palette tertiary, shadow=true]{Lemma 5}
    The executives face the following trade off when choosing $\bta$: higher tariffs decrease the probability that the trade agreement will be broken, but they also decrease welfare when the agreement is in force.
\end{beamerboxesrounded}
\end{frame}

\begin{frame}{Trade Agreements}

\begin{beamerboxesrounded}[upper=palette tertiary, shadow=true]{Result 4}
  The executives maximize their welfare by either (a) trading off reductions in the probability that the agreement will be broken with reductions in welfare under the agreement or (b) raising tariffs sufficiently high to ensure that the trade agreement will remain in force.
\end{beamerboxesrounded}
\end{frame}

\begin{comment}
\begin{frame}{Comparing Government Structures}
\begin{changemargin}{-1cm}{-1cm}
\begin{table}
\centering
  \begin{tabular}{lrrrr}
		\toprule
			Structure & $\bta$ & Lobbying & $B(\bta)$ & $\expect[\bta]$ \\
			\midrule
			Unitary, No Agreement & ---- & 0.0017 & 0 & 0.129 \\
			Unitary, Agreement & 0.118 & 0.0027 & 0 & 0.118 \\
			\\
			Exec-led Agreement (.25) & 0.078 & 0.0007 & 0.48 & 0.102 \\
			Exec-led Agreement (.00) & 0.106 & 0.0000 & 0 & 0.106 \\
			& {\color{gray}(.105)} & {\color{gray}(.0025)} & {\color{gray}(1)} & {\color{gray}(.129)} \\
		\bottomrule
	\end{tabular}
	\end{table}
\end{changemargin}
\end{frame}

\begin{frame}{Comparative Systems}
\begin{beamerboxesrounded}[upper=palette tertiary, shadow=true]{Conjecture}
    Lobbying effort and tariffs are reduced by the presence of executive branches who form a trade agreement.
\end{beamerboxesrounded}

\pause
\vskip.2in
\begin{beamerboxesrounded}[upper=palette tertiary, shadow=true]{Corollary}
    Tariffs are lower in countries with separated powers than in those with unitary governments.
\end{beamerboxesrounded}
\end{frame}
\end{comment}

\begin{frame}[label=Uncertainty]
\frametitle{Uncertainty}
\begin{itemize}[<+->]
	\item Added intuition for lobbying incentives and tariff-setting behavior
  \item High tariffs, low but positive lobbying effort
  \item Possibility of trade disruptions / ratification failure in equilibrium
  \item Different behavior at varying levels of uncertainty
\end{itemize}
\end{frame}

\section{Conclusion}
\subsection{}
\begin{frame}{Future Work}
\begin{itemize}[<+->]
	\item Comparative statics on weighting function
	\item Empirical Tests
	\item Parallel models for countries with different government structures
	\item Asymmetric agreements
	\item Repeated model
		\begin{itemize}
			\item Implications for design of dispute settlement procedures
		\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}{Conclusion}
Resolution of `Protection for Sale' Puzzle
\pause
\begin{itemize}[<+->]
	\item Non-unitary legislature: more realistic lobbying process
	\item Separation of Powers: tariffs set in shadow of lobbying
	\item Uncertainty: smoother results, additional intuition
\end{itemize}
\end{frame}

\begin{frame}{Conclusion}
\begin{itemize}[<+->]
	\item Social welfare maximizers ($a \approx \infty$) set high tariffs to achieve low lobbying expenditures (and low $\ga_{ML}$) 
	\item Executives use trade agreement to manipulate lobbying incentives
		\begin{itemize}
			\item Trade agreement as political commitment device
		\end{itemize}
	\item Firms lobby for trade war / ratification failure and achieve equilibrium trade disruptions in some cases
		\begin{itemize}
			\item Can examine the endogenous dynamics behind politically-driven failures to cooperate
		\end{itemize}
\end{itemize}

>>>>

\usepackage{cite}
\usepackage{xcolor,bm}
\usepackage{amsbsy,amssymb, amsmath, amsthm}
\usepackage{booktabs}
%David miller's fonts
	\usepackage[T1]{fontenc}
	\usepackage[boldsans]{ccfonts}
	\usepackage[euler-hat-accent]{eulervm}

@@

\section{Overview}
\subsection{Preview}
\begin{frame}
\frametitle{P}
In 
\pause
\begin{itemize}[<+->]
	\item r
  \item g
	\item a
\end{itemize}
\end{frame}

\begin{frame}{Preview of Results}
\begin{itemize}[<+->]
	\item H
		\begin{itemize}
			\item R
		\end{itemize}
	\item E
		\begin{itemize}
			\item P
			\item R
			\item T
		\end{itemize}
	\item T
\end{itemize}
\end{frame}

\begin{frame}{Outline of Talk}
\begin{enumerate}[<+->]
	\item B
	\item E
	\item R
	\item F
	\item Conclusion
\end{enumerate}
\end{frame}

\begin{comment}
\subsection{}
\begin{frame}
\frametitle{Related Literature}
\small Protection for Sale: Grossman $\&$ Helpman (1994)
\begin{itemize}
  \item \footnotesize Empirics: Goldberg $\&$ Maggi (1999), Gawande $\&$ Bandyopadhyay (2000), Mitra, Thomakos, $\&$  Ulubasoglu (2002)
  \item \footnotesize Mitra, Thomakos, $\&$  Ulubasoglu (2006), Bombardini (2008)
	\item \footnotesize Trade Wars and Trade Talks: Grossman $\&$ Helpman (1995)
\end{itemize}

@@

		%\item Autocracy vs. Democracy: Mitra, Thomakos and Ulubasoglu (2002), Aidt and Gassebner (2010)
\end{frame}
\end{comment}

@@

\vskip.2in
\begin{itemize}
	\item $e_b$: Lobbying effort for break decision
	\item $e_{tw}$: Lobbying effort for trade-war tariffs
	\item $(\tau^a,\tau^{*a})$: trade agreement tariffs
	\item $\tau^{\textit{ad}}$: tariff equivalent under anti-dumping duties
\end{itemize}

@@

\section{Results}
\subsection{Results: The Certain Case}
\begin{frame}{Timeline}
\begin{enumerate}
	\item \textbf{Formation}
		\begin{enumerate}[i.]
			\item Executives set trade policy in international agreement
		\end{enumerate}
	\item \textbf{Ratification / Maintenance}
		\begin{enumerate}[i.]
			\item Firms lobby legislatures to break agreement
			\item Legislatures decide to break or abide by agreement
		\end{enumerate}
	\item {\color{gray} \textbf{Trade War} (if agreement is broken)}
		\begin{enumerate}[i.]
			\item {\color{gray} Firms lobby legislatures to set high trade-war tariff}
			\item {\color{gray} Legislatures decide trade-war tariff}
		\end{enumerate}
	\item {\color{gray} Private actors make production, consumption decisions}
\end{enumerate}
\end{frame}

@@

\subsection{Full Results}
\begin{frame}{Political Uncertainty Illustration}
\pause
An Example (Bagwell $\&$ Staiger 2005)
\pause
\begin{itemize}[<+->]
	\item $D(P_i) = 1 - P_i$
	\item $Q_X(P_X) = \frac{P_X}{2}$, $Q_Y(P_Y) = P_Y$
	\item $P_X^W = \frac{4-3\tau}{7}$, $P_X = \frac{4+4\tau}{7}$
	\item $\ga(e,\ta) = 1.25 + e^{0.2} + \ta$ 
		\begin{itemize}
			\item $\ta \sim U[-0.25,0.25]$
		\end{itemize}
	\item $\ga_E = 1$
\end{itemize}

@@

\begin{frame}{Timeline}
\begin{enumerate}
	\item {\color{gray} \textbf{Formation}}
		\begin{enumerate}[i.]
			\item {\color{gray} Executives set trade policy in international agreement}
		\end{enumerate}
	\item \textbf{Ratification / Maintenance}
		\begin{enumerate}[i.]
			\item {\color{gray} Firms lobby legislatures to break agreement}
			\item Legislatures decide to break or abide by agreement
		\end{enumerate}
	\item {\color{gray} \textbf{Trade War} (if agreement is broken)}
		\begin{enumerate}[i.]
			\item {\color{gray} Firms lobby legislatures to set high trade-war tariff}
			\item {\color{gray} Legislatures decide trade-war tariff}
		\end{enumerate}
	\item {\color{gray} Private actors make production, consumption decisions}
\end{enumerate}
\end{frame}

@@

\section{Conclusion}
\subsection{}
\begin{frame}{Future Work}
\begin{itemize}[<+->]
	\item C
	\item E
	\item P
	\item A
\end{itemize}
\end{frame}

\begin{frame}{Conclusion}
\begin{itemize}[<+->]
	\item S
	\item E
	\item F
\end{itemize}
